Data Management for LSST Special Programs

### 2023-10-18 sec_sci.tex
minor updates

\section{Recommended Implementations to Enable Science from Special Programs}\label{sec:sci}

Recommendations regarding implementations by Rubin Data Management to meet the 
requirements related to Special Programs summarized in Section~\ref{sec:req},
which will enable science with data from Special Programs.

@@

\textbf{Program Labels -- } 
Based on survey cadeance or strategy.
Program labels might include, e.g., WFD, DDF, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Program labels should be propagated to single-visit processed and 
difference images and catalog sources.

It would not make sense to propagate program labels to deep coadds or 
object tables, as these data products could be a mix of programs.

The existence and propagation of these labels is not itself a requirement.
It is recommended for implementation for three reasons.

@@

The definitions of possible and necessary are ultimately left to the
discretion of Rubin Data Management in Operations.

@@

Specific examples are provided only to illustrate this recommendation,
not to place limits on what Special Processing might be done.
Note in particular that secondary science goals may be considered
to require User-Generated Processing.

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as "not necessary" and requiring 
User-Generated Processing.

\item \textbf{Necessary but not possible:}
in order to find the most distant, faint Kuiper Belt Objects in the DDF,
a specialized, computationally intensive form of "shift-and-stack" processing
is required for detection, but such algorithms are not used by the 
LSST Science Pipelines and so User-Generated Processing will be needed (Section~\ref{ssec:SPCS_TNO}).
A second example here is a twilight survey that uses non-standard visits 
which are outside the boundaries of what the LSST
Science Pipelines can process (Section~\ref{ssec:proc_bounds}).

@@

Rubin Observatory will reserve 10\% of its total data processing capacity for users.
This is already a requirement, as mentioned in Section~\ref{ssec:req_ug}.

This component would include {\it all} user processing and re-processing of any and 
all LSST data, including Special Programs.

@@

Users will need to be able to discover Special Programs data products when 
browsing data, such as all-sky maps.

>>>>

\section{Recommendations to Enable Science from Special Programs}\label{sec:sci}

Recommended implementations for Rubin Data Management to meet the 
requirements related to Special Programs (Section~\ref{sec:req})
and enable science with data from Special Programs.

@@

\textbf{Program Labels -- } 
Based on survey cadeance or strategy.
Program labels might include, e.g., WFD, DDF, TOO, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.
Program labels should be propagated to single-visit processed and 
difference images and catalog sources.
It would not make sense to propagate program labels to deep coadds or 
object tables, as these data products could be a mix of programs.

There are three main reasons for this recommendation.

@@

The definitions of possible and necessary, and the scope of Special Processing,
are ultimately left to the discretion of Rubin Data Management in Operations.
Note that secondary science goals may be considered as ``not necessary".

@@

Specific examples are provided only to illustrate this recommendation.
These examples do not place limits on what Special Processing might be done.

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.

\item \textbf{Necessary but not possible:}
in order to find the most distant, faint Kuiper Belt Objects in the DDF,
a specialized, computationally intensive form of ``shift-and-stack" processing
is required for detection, but such algorithms are not used by the 
LSST Science Pipelines and so User-Generated Processing will be needed (Section~\ref{ssec:SPCS_TNO}).
A second example here is a twilight survey that uses non-standard visits 
which are outside the boundaries of what the LSST
Science Pipelines can process (Section~\ref{ssec:proc_bounds}).

@@

Rubin Observatory will reserve 10\% of its total data processing capacity for users.
This is already a requirement, as mentioned in Section~\ref{ssec:req_ug}.
This component would include {\it all} user processing and re-processing of any and 
all LSST data, including Special Programs.

@@

Users will need to be able to discover Special Programs data products when 
browsing data.
This could be accomplished with, e.g., a toggle to overlay region boundaries
for Special Programs when browsing all-sky coadds built from WFD program data
in the Portal Aspect.
